---
title: "R/qtl: FR EtOH"
author: "Tiago da Silva Ribeiro"
date: '2022-07-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# French RILs - Ethanol tolerance mapping and epistasis test

This document contains the R/qtl code to analyze RIL data with the goal of finding QTLs and performing epistasis tests. Here, we are using FRxZI RILs and the phenotype analyzed is Ethanol Tolerance.

```{r etohload, echo=FALSE, warning=FALSE}
library(qtl)
library(ggplot2)
#setwd("Documents/git_repos/ril_epistasis/RILepi/rqtl/")
etoh = read.cross("csv", , "FR_etoh_rqtl.csv", sep=",", alleles=c("Z","F"), 
                  genotypes = c("FF","FZ","ZZ"))
```

#### Phenotype summary:
```{r etoh pheno sum, echo=FALSE, warning=FALSE, fig.height = 3, fig.width = 3, fig.align = "default"}
summary(etoh$pheno$etoh)
etohdf <- data.frame(etoh$pheno)
ggplot(etohdf, aes(x=RIL,y=etoh)) + geom_point() + theme_classic() + 
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggplot(etohdf, aes(x=etoh, y=..density..)) + geom_histogram(bins=20, colour="black", fill="light gray") + theme_classic() +
  geom_density(alpha=.2, fill="red") 
```

#### Genotype summary:
I am not doing the segregation distortion test, which has the default expectation of genotypes distributed following 1:2:1 fashion. The RILs would most likely have a 1:0:1 expectation, but we know that the heterozygous are not actually 0. That said, at a quick glance it seems like we have a biased towards having more FF than ZZ in the genome, so I will investigate that.
```{r geno expl, echo=FALSE, warning=TRUE, fig.height = 3, fig.width = 3, fig.align = "default"}
gt = geno.table(etoh)
gt[ gt$P.value < 1e-7, ]
```